Automating Verification of MPC Security

### 2023-06-06 popl24/metalang.tex
checkpoint

@@

\begin{lemma}[Input Encoding]
Assume given the following evaluation relation:
$$
\config{\prog_1}{E[\ttt{encode}(w,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If there exists gate identifiers $g_1, g_2$ such that:
\begin{enumerate}
  \item $g_1 \ne g_2 \wedge g_1,g_2 \not\in \gates(\prog_1)$
  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
  \item $\gNIMO(\prog_1,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1,v.\ttt{wv2},g_2)$
\end{enumerate}
then:
\begin{enumerate}
\item $g_1,g_2 \in  \gates(\prog_2)$
\item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
\item $$(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$$
\end{enumerate}
\end{lemma}

\begin{lemma}[Composition ($\eand$)]
Assume given the following evaluation relation:
$$
\config{\prog_1}{E[\ttt{andgg}(g,v_1,\ldots,v_5)]} \redxs \config{\prog_1;\prog_2}{E[v_6]}
$$
If there exist gate identifiers $g_1, g_2$ such that:
\begin{enumerate}
  \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2) \wedge \owl(\prog_1,v_3,g)$
  \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
\end{enumerate}
then:
\begin{enumerate}
  \item $g \in \gates(\prog_2)$
  \item $\gNIMO(\prog_1;\prog_2,v_6,g)$
  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}

>>>>

\subsection{Compositional Properties: Garbled $\NIMO$}

@@

\begin{definition}[Garbled $\NIMO$]
  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ iff for all $H$, $C$,
  both $w \in \{\ttt{p},\ttt{k}\}$, and all 
  memories $\store \in \mems(\iov(\prog) \cup \{ \itv \})$ where
  $
  \store(\itv) = \store(\flip{2}{g.w})$ or $\store(\itv) = \neg\store(\flip{2}{g.w})
  $
  it is the case that:
  $$
  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup \{\itv \}}})}{S_H} =
  \margd{(\condd{\progd(\prog,\be_w)}{\store_{S_C \cup V_C \cup \{\itv \}}})}{S_H}
  $$
\end{definition}

\begin{definition}
  We write $\owl(\prog, \{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
  iff for all $b$ and both $w \in \{\ttt{p},\ttt{k}\}$:
  $$
  (\margd{(\condd{\progd(\prog,\be_{w})}{\{ \flip{2}{g.w} \mapsto b \}})}{\itv})(\{ \itv \mapsto b\}) = 1
  $$
\end{definition}

\begin{lemma}[Input Encoding]
Assume given the following evaluation relation:
$$
\config{\prog_1}{E[\ttt{encode}(w,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If there exists gate identifiers $g_1, g_2$ such that:
\begin{enumerate}
  \item $g_1 \ne g_2 \wedge g_1,g_2 \not\in \gates(\prog_1)$
  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
\end{enumerate}
then:
\begin{enumerate}
\item $g_1,g_2 \in  \gates(\prog_2)$
\item $\gNIMO(\prog_1;\prog_2,v.\ttt{wv1},g_1) \wedge \gNIMO(\prog_1;\prog_2,v.\ttt{wv2},g_2)$
\item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}

\begin{lemma}[Composition ($\eand$)]
Assume given the following evaluation relation:
$$
\config{\prog_1}{E[\ttt{andgg}(g,v_1,\ldots,v_5)]} \redxs \config{\prog_1;\prog_2}{E[v_6]}
$$
If there exist gate identifiers $g_1, g_2$ such that:
\begin{enumerate}
  \item $g_1 \ne g_2 \ne g \wedge g_1,g_2 \in \gates(\prog_1) \wedge g \not\in \gates(\prog_1)$
  \item $\owl(\prog_1,v_1,g_1) \wedge \owl(\prog_1,v_2,g_2)$
  \item $\gNIMO(\prog_1,v_4,g_1) \wedge \gNIMO(\prog_1,v_5,g_2)$
\end{enumerate}
then:
\begin{enumerate}
  \item $g \in \gates(\prog_2) \wedge \owl(\prog_1;\prog_2,v_3,g)$
  \item $\gNIMO(\prog_1;\prog_2,v_6,g)$
  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}

### 2023-06-06 popl24/minicat.tex
checkpoint

\subsection{Syntax} The syntax of $\minifed$ includes a standard
boolean algebra with $\eand$, $\eor$, $\exor$, and $\enot$ as
primitives. In addition programs have different kinds of variables,
including \emph{secrets} $\secret{\cid}{w}$, \emph{flips}
$\flip{\cid}{w}$, and \emph{views} $\view{\cid}{w}$.  Each of these
variables are indexed by an ``owner'' client $\cid$ and distinguishing
string $w$. So for example, $\secret{\ttt{1}}{\ttt{foo}}$ is client
$\ttt{1}$'s secret called $\ttt{foo}$. All clients can make reference
to the \emph{oracle} through variables $\oracle{w}$- this is a form
of shared randomness that is standard in the MPC setting \cite{XXX}.
Each client can only compute on their own variables in
\emph{expressions} $\be$, and share values with other clients by
assignment to their views as in $\eassign{\view{\cid}{w}}{\be}$.  A
\emph{protocol} $\prog$ is a possibly empty sequence of view
assignments.

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips respectively. Given a program $\prog$, we
write $\iov(\prog)$ to denote the pair $(S,V)$ where $S$ is the set of
secret variables in $\prog$ and $V$ is the set of views in $\prog$,
and we write $\flips(\prog)$ to denote the set $F$ of flip variables
in $\prog$. For any set of variables $X$ and parties $P$, we write
$X_P$ to denote the subset of $X$ owned by any party in $P$, and we
write $X_H$ and $X_C$ to denote the subsets belonging to honest and
corrupt parties, respectively.

Expressions forms include the convenience
$\select{\be_1}{\be_2}{\be_3}$ which is like a conditional expression
in the binary field with:
$$
\select{\be_1}{\be_2}{\be_3} \equiv (\be_1\ \eand\ \be_2)\ \eor\ (\enot\ \be_1\ \eand\ \be_3)
$$
We also include \emph{oblivious transfer} $\OT{\be_1}{\be_2}{\be_3}$ as a primitive,
with semantics similar to $\ttt{select}$ but with important nuances
related to communication between clients as we discuss more below.
As we will demonstrate, it is not necessary to include $\ttt{OT}$
as a primitive since we can implement it in a provably secure
fashion. However its inclusion simplifies our presentation and is a useful
convenience.

@@

\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed}
$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
\end{array}
$$ 
\end{fpfig}

@@

Given a variable-free boolean expression $\be$, we write
$\cod{\be}$ to denote the standard interpretation of $\be$.
With the introduction of variables to expressions, we
have two concerns. First, we need to interpret expressions
with respect to a specific memory, and second, we need
to ensure that variables are used ``legally''. That is,
since expressions define local computation, all variables
used in an expression must belong to the same client.

\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& \etrue\\
\lcod{\store, \efalse}{\cid} &=& \efalse\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}

\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)

  \inferrule
  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
\end{mathpar}

>>>>

\subsection{Syntax} The syntax of $\minifed$, defined in
Figure \ref{fig-minifed-syntax}, includes a standard
boolean algebra with $\eand$, $\eor$, $\exor$, and $\enot$ as
primitives. In addition programs have different kinds of variables,
including \emph{secrets} $\secret{\cid}{w}$, \emph{flips}
$\flip{\cid}{w}$, and \emph{views} $\view{\cid}{w}$.  Each of these
variables are indexed by an ``owner'' client $\cid$ and distinguishing
string $w$. So for example, $\secret{\ttt{1}}{\ttt{foo}}$ is client
$\ttt{1}$'s secret called $\ttt{foo}$. All clients can make reference
to the \emph{oracle} through variables $\oracle{w}$- this is a form
of shared randomness that is standard in the MPC setting \cite{XXX}.
Each client can only compute on their own variables in
\emph{expressions} $\be$, and share values with other clients by
assignment to their views as in $\eassign{\view{\cid}{w}}{\be}$.  A
\emph{protocol} $\prog$ is a possibly empty sequence of view
assignments.

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips respectively. Given a program $\prog$, we
write $\iov(\prog)$ to denote $S \cup V$ where $S$ is the set of
secret variables in $\prog$ and $V$ is the set of views in $\prog$,
and we write $\flips(\prog)$ to denote the set $F$ of flip variables
in $\prog$. For any set of variables $X$ and parties $P$, we write
$X_P$ to denote the subset of $X$ owned by any party in $P$, and we
write $X_H$ and $X_C$ to denote the subsets belonging to honest and
corrupt parties, respectively.

Expressions forms include the convenience
$\select{\be_1}{\be_2}{\be_3}$ which is essentiall a conditional
expression with:
$$
\select{\be_1}{\be_2}{\be_3} \equiv (\be_1\ \eand\ \be_2)\ \eor\ (\enot\ \be_1\ \eand\ \be_3)
$$
We also include \emph{oblivious transfer} $\OT{\be_1}{\be_2}{\be_3}$ as a primitive,
with semantics similar to $\ttt{select}$ but with important nuances
related to communication between clients as we discuss more below.
As we will demonstrate, it is not necessary to include $\ttt{OT}$
as a primitive since we can implement it in a provably secure
fashion. However its inclusion simplifies our presentation and is a useful
convenience.

@@

\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
\end{array}
$$ 
\end{fpfig}

@@

Given a variable-free boolean expression $\be$, we write $\cod{\be}$
to denote the standard interpretation of $\be$.  With the introduction
of variables to expressions, we have two concerns. First, we need to
interpret variables with respect to a specific memory, and second,
we need to ensure that variables are used ``legally''. That is, since
expressions define local computation, all variables used in an
expression must belong to the same client.  Thus, we denote
interpretation of expressions $\be$ possibly containing variables as
$\lcod{\store,\be}{\cid}$, where $\store$ associates variables with
values and all variables must be owned by client $\cid$. This is
defined in Figure \ref{fig-minifed-interp}.

\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& \etrue\\
\lcod{\store, \efalse}{\cid} &=& \efalse\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}
\end{fpfig}

Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
This is defined in the obvious manner for view assignments other than through
$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
\end{mathpar}
For view assignments through $\ttt{OT}$, we define a different reduction rule that
captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
to the sender (through a view assignment), and only the selected bit is sent to the receiver:
\begin{mathpar}
  \inferrule
  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
\end{mathpar}
We define $\redxs$ as the reflexive, transitive closure of $\redx$.

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$.

\subsection{Examples}

\begin{verbatimtab}
v[2,0] := f[1,0];
v[0,0] := f[1,0] xor s[1,0];
v[2,1] := v[0,0];
v[2,2] := v[2,1] xor v[2,0]
\end{verbatimtab}

\begin{verbatimtab}
v[2, rd] := select[flip[2, d], H[r1], H[r0]];
v[1, e] := s[2, c] xor flip[2, d];
v[2, f0] := s[1, m0] xor select[v[1, e], H[r1], H[r0]];
v[2, f1] := s[1, m1] xor select[not v[1, e], H[r1], H[r0]];
v[2, mc] := (select[s[2, c], v[2, f1], v[2, f0]] xor v[2, rd])
\end{verbatimtab}
